---
title: "Final Project Template"
author: Dr. Petersen
output: pdf_document
---

## Briefly, check the logistic regression model assumptions

```{r}
df_logistic <- glm(diagnosis ~ radius_mean + texture_mean + concave points_mean + 
                               smoothness_mean + symmetry_mean, 
                   data = df,
                   family = binomial(link = "logit"))
summary(df_logistic)
```

## LINEARITY: The x's vs log odds are linear (monotone in probability)

```{r}
scatter.smooth(x = df$radius_mean, y = as.numeric(df$diagnosis) - 1)
scatter.smooth(x = df$texture_mean, y = as.numeric(df$diagnosis) - 1)
scatter.smooth(x = df$`concave points_mean`, y = as.numeric(df$diagnosis) - 1)
scatter.smooth(x = df$smoothness_mean, y = as.numeric(df$diagnosis) - 1)
scatter.smooth(x = df$symmetry_mean, y = as.numeric(df$diagnosis) - 1)
```

None of our plots show extreme non-linearity. We may assume the x's vs log odds are linear.

## INDEPENDENCE: The observations are independent
Each of our values in the dataset has a unique ID number. We will assume that our observations are independent.


## BERNOULLI: The response variable is Bernoulli
Our response variable is modeled by a Bernoulli distribution, with the only possible outcomes being Malignant (True) and Benign (False). 


## NO COUNFOUNDING VARIABLES: Additional predictor variables are not required
We will assume that we have accounted for all variables of significance in our model.


## INFLUENTIAL OBSERVATIONS: 
```{r}
plot(heart_logistic, which = 5)  # residuals vs leverage
```

None of our observations appear unduly influential based on our Cook's plot.

## MULTICOLLINEARITY: No multicollinearity, using VIFs
```{r}
vif(df_lm)
max(vif(df_lm))  # less than 10
mean(vif(df_lm))  # less than 5
```

Our Variance Inflation Factors (VIFs) have a mean less than 5 and a maximum less than 10. We may assume there is no extreme multicollinearity in our model.